Memories of migrations past: Sociality and cognition in dynamic, seasonal environments

### 2021-08-16 text/SocialMemoryMigration_FEE.tex
addressing comments

	While migration can be a very successful strategy, with migratory ecotypes of the same species often outnumbering non-migratory conspecifics.  Migratory caribou and reindeer \emph{Rangifer tarandus}, for example, are several orders of magnitude more abundant than non-migratory woodland, mountain and forest ecotypes \citep{FestaBianchet2011,Uboni2016}. However, the question of whether migratory animals are more or less resilient to environmental disruptions in the environment remains open and largely case-specific \citep{Moore2008, HardestyMoore2017, Xu2021}. On the one hand, migratory species may be more vulnerable as disruptions in either of the seasonal ranges or along a migratory corridor can have significant negative impacts \citep{Wilcove2008, Seebacher2015, Kauffman2021}. On the other hand, migratory species might be more resilient due to their general wide-ranging mobility \citep{Robinson2009}. The resilience of a migratory populations depends on the plasticity and adaptability of a population, which can take multiple forms, reflecting variation in \textit{where}, \textit{when} and \textit{whether} the migration occurs \citep{Gurarie2017, Xu2021}.

@@

	\noindent where each of the six parameters is updated according to equaton \ref{eq_memory} identically. The estimates $\widehat{\theta_y}$ are obtained via a least-squares minimization of the migration track ($m(t,\theta) = \int_0^t v_m(t',\theta_y) \, \Delta t'$) against the spatial mean of the population process in year $y$ (i.e.~$\widehat{u}(t) = \int_X u_y(t, x) dx$). The parameter $\kappa \in (0,1)$ captures the reliance on that long-term memory. When $\kappa = 0$, all of the actionable memory is from the preceding year. When $\kappa = 1$, the actionable memory is entirely the reference memory.

@@

	We ran this model on a spatial domain $x \in [-100,100]$, and a periodicity $\tau = 100$ (i.e.~100 day year~s). We were interested in an approximately periodic resource dynamic, i.e.~one in which $h(x,t \approx h(x, t-\tau))$. We generated two types of resource distributions. A ``non-surfable'' resource (\emph{island resource}), and weakly surfable resource (\emph{drifting resource}). Both are characterized by a peak in time and space centered at $m_x$ at $m_t$, and $-m_x$ at $\tau - m_t$ (for example, locations 30 and -30 at times 25 and 75, respectively). These pulses have a shared time scale of duration $s_t$ and a spatial scale of extent $s_x$, the standard deviation in the time and space dimension respectively. The island resource is simply two uncorrelated bivariate normal
	distributions 
	$$h(x,t) = K\,(\Phi(m_x, m_t, s_x, s_t) + \Phi(-m_x, \tau - m_t, s_x, s_t))$$ 
	\noindent where $\Phi$ is the bivariate Gaussian distribution function, and the normalizing constant $K$ is selected such that the average total amount of resource throughout the year is 1, i.e.~$\frac{1}{\tau} \int_T\int_X h(x,t) dx\,dt = 1$.

@@

	Within a given year, the resource is entirely symmetric: $h_y(x,t) = h_y(-x, \tau-t)$. However, in scenarios exploring climate change we allow the peaks to vary with directional trend and stochasticity according to: $m_x(y) \sim {N}(\mu_x + \gamma_x\,y, \sigma_x)$ and $m_t(y) \sim {N}(\mu_t + \gamma_t\,y, \sigma_t)$, where the $\mu$, $\gamma$ and $\sigma$ terms are the mean, slope and variance, respectively, for the location and time duration of the pulse. Thus, if $\gamma=0$ and $\sigma=0$, the conditions are constant across years and if $\gamma_x > 0$ there is a shift of the resource towards the extremes of the domain. While we did note explore phenological shifts in timing, those can readily be modeled as well. These trends model the pole-ward shift of peak resources and the earlier spring phenology occurring with a warming global climate \citep{Renner2018}. The spatial and temporal scales of the resource peak ($s_x$ and $s_t$) remain constant in all of our simulations.

@@

	As anticipated, better adaptation to the drifting resource correlated strongly with higher foraging efficiency (inset boxplots).

@@

		\caption{ \label{fig_trendstochasticity} Role of reference memory in adapting to climate change for increasingly stochastic resource dynamics. We ran the model with a moderiate rate of climate change (mean shift: 0.5/year) at five increasing levels of stochasticity (inter-annual standard deviation of resource peak 0, 3, 6 and 12, left to right panels). For non-zero stochasticity, we ran the process 30 times and present the mean and standard error of the spatial adaptation index across various values of the reference memory parameter $\kappa$: where $\kappa = 0$, the system modifies its migration based entirely on recent experiences; at $\kappa = 1$, the memory never changes from the reference memory. Other parameter values are resource following $\alpha = 100$, social attraction $\beta = 400$, and diffusion $\varepsilon = 4$.}

@@

	By allowing a population to adjust its migratory behavior based on recent experiences with the resource location, our model captured many fundamental phenomena related to migration. The model was able to emulate the successful navigation of an environment with temporally and spatially isolated seasonal resource patches, the spontaneous emergence of a migratory behavior, and intrinsic robustness to changes in those environmental resources, whether those changes were steadily shifting trends or inter-annual stochasticity.  In a generalized way, it captured all three fundamental types of migration plasticity: the \textit{where}, the \textit{when} and the \textit{whether} \citep{Xu2021}.

@@

	\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
	%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
	\bibliography{memorymigration.bib}

>>>>

	While migration can be a very successful strategy, with migratory ecotypes of the same species often outnumbering non-migratory conspecifics.  Migratory caribou and reindeer \emph{Rangifer tarandus}, for example, are several orders of magnitude more abundant than non-migratory woodland, mountain and forest ecotypes \citep{FestaBianchet2011,Uboni2016}. However, the question of whether migratory animals are more or less resilient to environmental disruptions in the environment remains open and largely case-specific \citep{Moore2008, HardestyMoore2018, Xu2021}. On the one hand, migratory species may be more vulnerable as disruptions in either of the seasonal ranges or along a migratory corridor can have significant negative impacts \citep{Wilcove2008, Seebacher2015, Kauffman2021}. On the other hand, migratory species might be more resilient due to their general wide-ranging mobility \citep{Robinson2009}. The resilience of a migratory populations depends on the plasticity and adaptability of a population, which can take multiple forms, reflecting variation in \textit{where}, \textit{when} and \textit{whether} the migration occurs \citep{Gurarie2017, Xu2021}.

@@

	\noindent where each of the six parameters is updated according to equaton \ref{eq_memory} identically. The estimates $\widehat{\theta_y}$ are obtained via a least-squares minimization of the migration track ($m(t,\theta) = \int_0^t v_m(t',\theta_y) \, dt'$) against the spatial mean of the population process in year $y$ (i.e.~$\widehat{u}(t) = \int_X u_y(t, x) dx$). The parameter $\kappa \in (0,1)$ captures the reliance on that long-term memory. When $\kappa = 0$, all of the actionable memory is from the preceding year. When $\kappa = 1$, the actionable memory is entirely the reference memory.

@@

	We ran this model on a spatial domain $x \in [-100,100]$, and a periodicity $\tau = 100$ (i.e.~100 day years). We were interested in an approximately periodic resource dynamic, i.e.~one in which $h(x,t) \approx h(x, t-\tau))$. We generated two types of resource distributions. A ``non-surfable'' resource (\emph{island resource}), and weakly surfable resource (\emph{drifting resource}). Both are characterized by a peak in time and space centered at $m_x$ at $m_t$, and $-m_x$ at $\tau - m_t$ (for example, locations 30 and $-30$ at times 25 and 75, respectively). These pulses have a shared time scale of duration $s_t$ and a spatial scale of extent $s_x$, the standard deviation in the time and space dimension respectively. The island resource is simply two uncorrelated bivariate normal
	distributions

	$$h(x,t) = K\,(\Phi(m_x, m_t, s_x, s_t) + \Phi(-m_x, \tau - m_t, s_x, s_t))$$

	\noindent where $\Phi$ is the bivariate Gaussian distribution function, and the normalizing constant $K$ is selected such that the average total amount of resource throughout the year is 1, i.e.~$\frac{1}{\tau} \int_T\int_X h(x,t) dx\,dt = 1$.

@@

	Within a given year, the resource is entirely symmetric: $h_y(x,t) = h_y(-x, \tau-t)$. However, in scenarios exploring climate change we allow the peaks to vary with directional trend and stochasticity according to: $m_x(y) \sim {N}(\mu_x + \gamma_x\,y, \sigma_x)$ and $m_t(y) \sim {N}(\mu_t + \gamma_t\,y, \sigma_t)$, where the $\mu$, $\gamma$ and $\sigma$ terms are the mean, slope and variance, respectively, for the location and time duration of the pulse. Thus, if $\gamma=0$ and $\sigma=0$, the conditions are constant across years and if $\gamma_x > 0$ there is a shift of the resource towards the extremes of the domain. While we did not explore phenological shifts in timing, those can readily be modeled as well. These trends model the pole-ward shift of peak resources and the earlier spring phenology occurring with a warming global climate \citep{Renner2018}. The spatial and temporal scales of the resource peak ($s_x$ and $s_t$) remain constant in all of our simulations.

@@

	As anticipated, better adaptation to the drifting resource correlated strongly with higher foraging efficiency (inset boxplots).\\

@@

		\caption{ \label{fig_trendstochasticity} Role of reference memory in adapting to climate change for increasingly stochastic resource dynamics. We ran the model with a moderate rate of climate change (mean shift: 0.5/year) at five increasing levels of stochasticity (inter-annual standard deviation of resource peak 0, 3, 6 and 12, left to right panels). For non-zero stochasticity, we ran the process 30 times and present the mean and standard error of the spatial adaptation index across various values of the reference memory parameter $\kappa$: where $\kappa = 0$, the system modifies its migration based entirely on recent experiences; at $\kappa = 1$, the memory never changes from the reference memory. Other parameter values are resource following $\alpha = 100$, social attraction $\beta = 400$, and diffusion $\varepsilon = 4$.}

@@

	By allowing a population to adjust its migratory behavior based on recent experiences with the resource location, our model captured many fundamental phenomena related to migration. The model was able to emulate the successful navigation of an environment with temporally and spatially isolated seasonal resource patches, the emergence of a migratory behavior from an essentially resident or naive initial condition, and intrinsic robustness to changes in those environmental resources, whether those changes were steadily shifting trends or inter-annual stochasticity.  In a generalized way, it captured all three fundamental types of migration plasticity: the \textit{where}, the \textit{when} and the \textit{whether} \citep{Xu2021}.

@@

	\bibliographystyle{frontiersinSCNS_ENG_HUMS} 
	\bibliography{memorymigration.bib}
